---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# kapa

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![CRAN status](https://www.r-pkg.org/badges/version/kapa)](https://CRAN.R-project.org/package=kapa)
<!-- badges: end -->

The kapa R package enables creating [ellmer](https://ellmer.tidyverse.org/) tools from [kapa.ai](kapa.ai) projects. kapa.ai projects compile context from various resources and surface that information through search.

## Installation

You can install the development version of kapa like so:

``` r
pak::pak("simonpcouch/kapa")
```

Once the package is installed, you will need a kapa.ai API key. Set this to the `KAPA_API_KEY` environment variable.

## Example

`kapa_search_tool()` returns an ellmer tool for the provided kapa.ai `project`. `kapa_search_tool(project)` defaults to `Sys.getenv("KAPA_PROJECT_ID"`)—I've set mine to [a project](https://demo.kapa.ai/widget/posit) that compiles documentation and GitHub history on a few tidyverse packages:

```{r}
library(kapa)

tidy_search <- kapa_search_tool()

tidy_search
```

I can add this tool to any ellmer chat object to allow the chat to search this kapa.ai project before responding to questions. For example, `coord_flip()` was deprecated after Claude Sonnet 4 was trained, but using the kapa.ai search allows the model to discover that the function was deprecated before responding to the user:

```{r}
#| eval: false
library(ellmer)

client <- chat_anthropic(
  "Use the provided tool to research your solution before responding tersely.",
  model = "claude-sonnet-4-20250514"
)
client$register_tool(tidy_search)

client$chat(
  "
Could you flip the axes in this plotting code?

ggplot(diamonds, aes(x = carat)) +
  geom_histogram() +
  scale_x_reverse()
"
)
#> ◯ [tool call] Posit_kapa_search(query = "ggplot flip axes coord_flip
#> histogram x y axis")
#> ● #> {"search_results":[{"content":"# ggplot2 > Reference\n## Cartesian …
#> To flip the axes in your plotting code, you have two main options:
#>
#> **Option 1: Using `coord_flip()` (keep your current code structure)**
#> ```r
#> ggplot(diamonds, aes(x = carat)) +
#>   geom_histogram() +
#>   coord_flip() +
#>   scale_x_reverse()
#> ```
#>
#> **Option 2: Swap the aesthetics (preferred modern approach)**
#> ```r
#> ggplot(diamonds, aes(y = carat)) +
#>   geom_histogram() +
#>   scale_y_reverse()
#> ```
#>
#> The second option is preferred because `coord_flip()` is now superseded. When
#> you map the variable to the `y` aesthetic instead of `x`, the histogram will
#> automatically be oriented horizontally, and you use `scale_y_reverse()`
#> instead of `scale_x_reverse()` to maintain the reverse scaling.
```

To start an MCP server with this tool, use []`mcptools::mcp_server()`](https://posit-dev.github.io/mcptools/) with `tools = list(<your_tool>)`.
